Mobile Assets in Semantic Digital Twins

### 2023-05-12 paper/main.tex
Add second draft of implementation

Furthermore, there is existing research about using Building Information Modeling (BIM) with standard data formats, such as JSON\footnote{\url{https://www.json.org/json-en.html}} or Resource Description Framework (RDF)\footnote{\url{https://www.w3.org/RDF/}} to better define static building infrastructure \cite{pauwels_live_2023}. \citeauthor{pauwels_live_2023} proposes how, and to what extent, such data can be used in robot navigation. Although BIM can be used, by default it has some challenges regarding interoperability due to heterogeneous data \cite{dinis_bim_2022,godager_concept_2021}.

@@

@@

As of today, there is no agreed-upon definition of a DT. However, it can be referred to as a digital representation of some physical system. There is no restriction on the size or scale of the system, which means the entity could be a single asset or a system of systems for that matter \cite{li_digital_2022, waszak_let_2022}. The  counterpart existing in real life is commonly referred to as the physical twin (PT), and is observed by the DT in near real-time. There should be a bidirectional flow of data between the DT and the PT throughout the life cycle of the PT. In other words, the DT gets updated sensor data from the PT, and sends informed decisions back to it \cite{madni_leveraging_2019, waszak_let_2022, kamburjan_digital_2022}.

There are multiple misconceptions about DTs. One such misconception is that a digital shadow (DS) is some sort of a DT. This misconception often leads people to building shadows instead of DTs. \cite{fuller_digital_2020, li_digital_2022}.  Although very similar, a digital shadow (DS) does not have the bidirectional data flow that a DT does. Instead, there is a one-way flow from the PT to the digital counterpart, which often leads to a change in the DS, but not vice versa \cite{kritzinger_digital_2018, li_digital_2022}. Secondly, as told by \citeauthor{fuller_digital_2020}, a common misconception is that the DT must be an exact 3D copy of a physical entity, such as a building.

In order to capture diverse data and formalise knowledge represented in DTs, knowledge graphs should be used \cite{kamburjan_programming_2021, waszak_let_2022}. \citeauthor{waszak_let_2022} created a DT architecture in which static and dynamic data sources were not explicitly stored in a common knowledge graph, but instead linked to one another for better scaling and heterogeneity.

DTs have had many applications in various disciplines over the years. Although the idea of a digital twin was adopted early in the aerospace industry by NASA \cite{li_digital_2022}, later years have seen widespread use elsewhere \cite{fuller_digital_2020, waszak_let_2022, macchi_exploring_2018}:

@@

There are many challenges with DTs. Some of these concern data, privacy, and costs of modeling physical assets \cite{fuller_digital_2020, waszak_let_2022}. DTs depend on the accuracy of the sensor data sent by the PTs. General Data Protection Regulation (GDPR)\footnote{\url{https://gdpr-info.eu/}} works to ensure privacy and security of personal data. As an example, no personal data should be collected without the permission of the \emph{data subject} (i.e. user), stated in GDPR Article 13 and 14 \cite{noauthor_guide_nodate}.

It is important to understand what a DT is and what it is not. We should try to "close the loop" by sending informed decisions back to the physical counterpart so that it reflects the newest state of the DT. A DS could be used alongside the DT with the only goal of being a timestamped snapshot of some specific physical asset \cite{bergs_concept_2021}. When implementing one should also think about the challenges of digital twins, such as the accuracy of sensor data, privacy concerns, and costs. We should not get too caught up with modeling an exact 3D replica of an asset. On the other hand, we \emph{should} use existing semantic technologies such as knowledge graphs for formal knowledge representation from a variety of data sources.

\subsection{Semantic Web}
Semantic Web was \emph{envisioned} by Sir Tim Berners-Lee as an extension of the World Wide Web (WWW) in which data was linked. Linked data is about making links so people and machines can explore the web of data, and should follow the four rules \cite{tim_berners-lee_linked_nodate} below:

@@

    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.2]{graphics/rdf_triple_simple.png}
        \caption{Example of an RDF triple consisting of a subject, a predicate, and an object.}
        \label{fig:rdf_triple_simple}
    \end{figure}

@@

\begin{itemize}
    \item \textbf{Web Ontology Language (OWL)} is a language for expressing rich knowledge of concepts (phrases in natural language) in an application domain \cite{szolovits_overview_1977}, as well as relations between entities. More specifically, ontologies describe domains in terms of classes, properties and individuals \cite{bechhofer_owl_2009}.

@@

There also exist other technologies that provide reasoner tools. Some notable mentions include, but are not limited to:
\begin{itemize}
    \item{\textbf{Apache Jena}} is an open source framework that has the Inference API which include reasoner tools that can be used to reason over data, as well as checking content of triple stores \cite{noauthor_apache_nodate-1}. The source code is publicly available\footnote{\url{https://github.com/apache/jena}}.
    \item{\textbf{OWL API}} is a Java framework that let us create, serialise or manipulate ontologies written in OWL \cite{noauthor_owl_nodate}, and its source code is openly accessible\footnote{\url{https://github.com/owlcs/owlapi}}. Its recent development is more focused towards OWL 2 however.
    \item\hyperref[subsec:SMOL]{\textbf{SMOL}} provide reasoning tools through its interpreter, which implements \emph{semantic lifting} (i.e. generating a knowledge graph from the current state of a SMOL program) \cite{noauthor_semantic_nodate}.

@@

\subsection{Databases}\label{subsec:Databases}
Databases are collections of information that exist for a long time, often over many years \cite{garcia-molina_database_2002}. They can be divided into relational and non-relational (NoSQL) databases and are often compared as such \cite{mohamed_relational_2014}. Relational databases have existed for many decades and store data in tables (which consist of columns and rows), whilst \emph{NoSQL} (i.e. not only SQL) databases store data in JSON documents \cite{garcia-molina_database_2002, sudiartha_data_2020}. Examples of relational databases are MySQL and PostgreSQL, whereas examples of NoSQL databases are: 
\begin{itemize}
    \item \textbf{MongoDB} is a well-known NoSQL database for storing data in JSON documents and its source code is publicly available\footnote{\url{https://github.com/mongodb/mongo}}.
    \item \textbf{Firebase Realtime Database} is a cloud-hosted NoSQL database that stores data as JSON  \cite{noauthor_firebase_nodate}. Furthermore, data is stored and synchronized in real-time to cross-platform clients.
    \item \textbf{InfluxDB} is a NoSQL database for storing \emph{time series data} (i.e. "sequence of data points indexed in time order" \cite{noauthor_what_nodate})  \cite{noauthor_influxdb_nodate}, and its source code is also openly accessible\footnote{\url{https://github.com/influxdata/influxdb}}.
\end{itemize}

@@

\newpage
\section{Problem analysis and design}\label{sec:Analysis}
This chapter examines the problem statement (H) and its three guiding research questions (RQ1, RQ2, RQ3) in \ref{sec:Introduction}(Introduction), based on the theoretical background information and general code examples in \ref{sec:Background}(Background). We create a building domain based on the analysis of the research questions. Then we examine what a dynamic asset model entails, and how static data can be separated from dynamic data in it. Different technologies described in \ref{sec:Background}, such as databases, frameworks for creating applications (apps), and semantic digital twins are considered. Lastly, we present the resulting structural requirements as the basis for our proof-of-concept implementation.

@@

\subsubsection{Second research question (RQ2)}
RQ2 asks whether we can "enable bidirectional data flow between the PT and DT, such that the digital twin gets updated sensor data from the mobile devices in the PT and sends informed decisions back". Bidirectional data flow is data that is exchanged both ways, and not just in one direction between the physical twin (PT) and digital twin (DT), or vice versa. The DT observes the PT in near real-time and gets updated sensor data from its physical counterpart. The DT should then send informed decisions back to it. A lot of value is derived from the DT making an informed decision on behalf of a mobile asset in the PT, and it is crucial that the appropriate smartphone user is notified.

@@

\subsection{Building domain}
From considering the three research questions, we have created a possible domain including a built environment based on static data, as well as movable entities with their physical location as dynamic data, as shown in Figure \ref{fig:static_built_environment}. \textbf{Physical 2D space} is provided for context of outside and is in two dimensions as described in \ref{subsec:Scope}(Thesis scope and outline). This means that altitude is not taken into account. Note that the relationships between some entities use \emph{composition} (i.e. indicates a very strong relationship between the entities, which means that if the owing entity (e.g. \textbf{Physical 2D space}) is destroyed, so is the entity linked to it (e.g. \textbf{Building} \cite{pilone_uml_2005}) (In this case the composition can be thought of as a black hole..)

@@

It is important to look at the differences between the sensor data and the informed decisions in terms of what the data can be used for. The DT gets updated sensor data from the PT, and sends informed decisions back, as RQ2 states. The updated sensor data from the PT is the newest physical location from the mobile assets in the PT. It is crucial that the PT can use this data \emph{as is}, and therefore the data should be exact coordinates (latitude, longitude), and identifiable. We should also think about what an informed decision means. An informed decision is a conclusion made by the DT when there is sufficient information to determine something on behalf of the PT to control it. An example of this is the DT sending a warning back to the PT when it is determined that smartphones are inside a critical area. The analysis of what a critical area \emph{is} however, is not taken into account, as stated in \ref{subsec:Scope}(Thesis scope and outline). Therefore, the DT should not decide that an area is dangerous. This could be a job for an operator responsible for defining the areas in the asset model.

@@

The \textbf{Server} in Figure \ref{fig:initial_components} has not been explicitly described yet in this thesis as its main function was to enable a bidirectional flow of data between the PT and the digital twin (DT). Most people in general are familiar with Java,  a well-known and established programming language, and our fellow students are quite familiar with its syntax as well.

\subsubsection{Semantic Digital Twins}\label{subsubsection:SemanticDigitalTwins}
\ref{subsec:SMOL}(SMOL) describes that SMOL (Semantic Micro Object Language) can be used as a framework for creating digital twins (DTs) and that it integrates semantic technologies. Knowledge graphs can be generated by using the \verb|dump|-command in the REPL provided by the SMOL interpreter. The knowledge graph can capture a knowledge base, such as a dynamic asset model, in which static data is separated from dynamic data. This asset model combined with the SMOL program, produces the knowledge graph. It is also possible to access the asset model from the SMOL program. As SMOL also integrates semantic languages, such as SHACL for validation and SPARQL for querying, it could be a good fit for our implementation.

@@

\noindent\textbf{Structural requirements:}
\begin{itemize}
    \item \textbf{Client}
    \begin{itemize}
        \item Client can choose to send its newest status (ID, latitude, longitude) 
        \item Appropriate endangered clients needs to be warned 
        \item Show appropriate in-app messages to the user
        \item Ask for permission to track the physical position of the device
        \item Clearly show when the physical location of the device is tracked
        \item Clearly show when the physical location of the device is \emph{not} tracked
    \end{itemize}
    \item \textbf{Server}
    \begin{itemize}
        \item Reloads asset model with tracked clients
        \item Access and read the generated knowledge graph
        \item \textbf{1.} Forward the sensor data to a time series database, which the DT accesses
        \item \textbf{2.} Forward the informed decisions to a real-time database, which the PT accesses
        \item Should be possible to do \textbf{1.} and \textbf{2.} above at the same time
    \end{itemize}
    \item \textbf{Databases}
    \begin{itemize}
        \item Realtime database
        \begin{itemize}
            \item Store the ID and physical location of the clients
            \item Store data about which clients (in the PT) the DT made an informed decision on behalf of
        \end{itemize}
        \item Time series database
        \begin{itemize}
            \item Store the time series data
        \end{itemize}
    \end{itemize}
    \item \textbf{Digital Twin}
    \begin{itemize} 
        \item \textbf{3.} DS gets the newest sensor data from the time series database
        \item \textbf{4.} Access the smartphones by their IDs in the asset model
        \item Create smartphone objects by combining data from \textbf{3.} and \textbf{4.} above. 
        \item Access \emph{only} the areas from the asset model that are \emph{critical}
        \item Check if smartphones are inside any critical areas
        \item Handle if there are any smartphones that don't have a registered physical location
        \item Provide examples of SPARQL queries
        \item Show informative messages in the REPL during program execution
    \end{itemize}
\end{itemize}

@@

\section{Implementation}\label{sec:Implementation}
This chapter describes the implemented solution based on the structural requirements in \ref{subsec:Requirements}, by describing each of the following parts in it:
\begin{itemize}
    \item Client
    \item Server
    \item Databases
    \item Digital Twin
\end{itemize}
In the first section we describe an \emph{early version} of the final implementation based on the initial architecture of software components in Figure \ref{fig:initial_components}.

Then we look at a revised version in Figure \ref{fig:components}, and describe what we implemented in the final solution. For each of the architectural components in the figure, we also reference the section that describes it in \ref{sec:Background}(Background), and the design decision that enables it in \ref{sec:Analysis}(Analysis and design). We also reference the corresponding \emph{item} in the structural requirement \emph{part} we are describing.

\subsection{Early implementation}\label{subsec:EarlyImplementation}
Due to the extensive architecture, we chose to implement an early version by letting all of the architectural components in Figure \ref{fig:initial_components} communicate first. The \textbf{Physical Twin} in the figure consists of \textbf{Mobile Assets} which are smartphones. Based on the background information in \ref{subsec:AppDevelopment}(Background - App Development), and from further narrowing of the technologies in \ref{subsubsection:AppDevelopment}(Analysis - App Development), a Flutter app should be created for cross-platform support.

Figure \ref{fig:recording_off} and \ref{fig:recording_on} show the location recording toggle function in the app, off and on, respectively. The recording is off by default, and when the button is pressed, it starts tracking the position of the device and changes color. Then when it is pressed again, the latest physical location of the device is sent, alongside its ID, as a status message to the server.

@@

We created a Maven project (a tool for managing the project) and used the object-oriented programming language Java for the server implementation. To let the server track many clients, we first assigned an ID to the client when it connected to the server (the server kept track of all clients by their IDs), which was then later sent alongside the client's physical location as status update messages. As this was an early implementation, we implemented client-server communication over TCP (Transmission Control Protocol). This is also further discussed in \ref{subsec:ClientServerCommunication}(Client-server communication).

An asset model (ontology) of a simple building was also created as shown in Figure \ref{fig:simple_building} (this time with coordinates as data properties which were "added" to individuals of the \verb|MovableEntity| class suggested in \ref{subsubsec:DynamicAssetModel}(Dynamic asset model). We used Protégé to create it and export it as an OWL document (file). This file was then accessed in a new SMOL program based on the background information in \ref{subsubsec:Technologies}(Technologies) and \ref{subsubsec:CoreLanguage}, and from further analysis in \ref{subsubsec:StaticAndDynamicData}(Static and dynamic data).

@@

\subsection{Implemented solution}\label{subsec:ImplementedSolution}
Figure \ref{fig:components} shows the architectural components in the proof-of-concept implementation. More specifically, it illustrates the \textbf{Mobile Assets} in the \textbf{Physical Twin}, the \textbf{Digital Twin}, and the \textbf{Server} which enables communication between them both ways. In addition to this, the figure illustrates that \textbf{Firebase Realtime database} is used as a channel between the \textbf{Physical Twin} and the \textbf{Server}, and that \textbf{InfluxDB time-series database} is used for storing the sensor data which \emph{can} be accessed from the \textbf{SMOL} program, as shown in the example function in \ref{fig:access_time_series}.

@@

For further reference on how \emph{we} set up SMOL for the this implementation, see \ref{subsec:Setup}(Setup).

In the following sections, we are describing the architectural \emph{components} one by one based on the background information and the analysis, and by referencing the corresponding \emph{part} (Client, Server, Databases, or Digital Twin) in \ref{subsec:Requirements}{Structural requirements}.

\subsubsection{Client}
\subsection{Digital Twin}\label{subsec:DigitalTwins}
A Flutter app written in Dart was created, based on the background information on app development in industry in \ref{subsec:AppDevelopment}(Background - App Development), and from the analysis of technologies in \ref{subsubsection:AppDevelopment} which indicates that Flutter is popular in industry. \ref{subsec:EarlyImplementation}(Early implementation) describes the toggling of location recording in the app, but from the paragraph about GDPR (General Data Protection Regulation) in \ref{subsec:DigitalTwins} and the requirement "Ask for permission to track the physical position of the device" we implemented asking for permission to access it. The section \ref{subsec:privacy} discusses privacy further.

The source code of the app was structured into the parts: 
\begin{itemize}
    \item \verb|main.dart| which initializes the app and Firebase
    \item \verb|map.dart| which is the UI (User interface) in the app. This is done by rendering Flutter widgets which behave much like components in e.g. React. Simply put, the UI is what the user \emph{sees} in the app, and the widgets are parts of what the user \emph{sees}.
    \item \verb|service.dart| communicates (reads, writs) with Firebase
    \item the model-class \verb|status.dart| can be used to create status objects sent to Firebase
\end{itemize}
We used the Google Maps SDK (Software Development Kit) in the implementation to integrate the map and its functions in the app, as described in \ref{subsec:AppDevelopment}(App Development).

Some of the requirements set forth in the structural requirements were already implemented in the earlier version of the app, such as "Client can choose to send its newest status (ID, latitude, longitude)" However, there are two requirements stating that we have to \emph{clearly} show when the recording function is off/on. In order to clearly show it for those that are color blind as well, the colors of the button changed to blue and red, respectively. Figure \ref{fig:color_theme} shows the final color theme, which is located in the previously mentioned file \verb|main.dart|:

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.70]{graphics/color_theme.png}
    \caption{The color theme in the app.}
    \label{fig:color_theme}
\end{figure}

>>>>

Furthermore, there is existing research about using Building Information Modeling (BIM) with standard data formats, such as JSON or Resource Description Framework (RDF), to better define static building infrastructure \cite{pauwels_live_2023}. \citeauthor{pauwels_live_2023} proposes how, and to what extent, such data can be used in robot navigation. Although BIM can be used, by default it has some challenges regarding interoperability due to heterogeneous data \cite{dinis_bim_2022,godager_concept_2021}.

@@

We are also assuming the PT contains a building in which there can be smartphones since we create a digital representation of a building in the DT.

@@

As of today, there is no agreed-upon definition of a DT. However, it can be referred to as a digital representation of some physical system. There is no restriction on the size or scale of the physical system, which means the entity could be a single asset or a system of systems for that matter \cite{li_digital_2022, waszak_let_2022}. It exists in real life and is commonly referred to as the physical twin (PT), and is observed by the DT in near real-time. There should be a bidirectional flow of data between the DT and the PT throughout the life cycle of the PT. In other words, the DT gets updated sensor data from the PT, and sends informed decisions back to it \cite{madni_leveraging_2019, waszak_let_2022, kamburjan_digital_2022}.

There are multiple misconceptions about DTs. One such misconception is that a digital shadow (DS) is some sort of a DT. This misconception often leads people to build shadows instead of DTs. \cite{fuller_digital_2020, li_digital_2022}.  Although very similar, a digital shadow (DS) does not have the bidirectional data flow that a DT does. Instead, there is a one-way flow from the PT to the digital counterpart, which often leads to a change in the DS, but not vice versa \cite{kritzinger_digital_2018, li_digital_2022}. Secondly, as told by \citeauthor{fuller_digital_2020}, a common misconception is that the DT must be an exact 3D copy of a physical entity, such as a building.

In order to capture diverse data and formalize knowledge represented in DTs, knowledge graphs should be used \cite{kamburjan_programming_2021, waszak_let_2022}. \citeauthor{waszak_let_2022} created a DT architecture in which static and dynamic data sources were not explicitly stored in a common knowledge graph but instead linked to one another for better scaling and heterogeneity.

DTs have had many applications in various disciplines over the years. Although the idea of a digital twin was adopted early in the aerospace industry by NASA \cite{li_digital_2022}, later years have seen widespread use elsewhere \cite{fuller_digital_2020, waszak_let_2022, macchi_exploring_2018}, such as:

@@

There are many challenges with DTs. Some of these concern data, privacy, and costs of modeling physical assets \cite{fuller_digital_2020, waszak_let_2022}. General Data Protection Regulation (GDPR)\footnote{\url{https://gdpr-info.eu/}} works to ensure the privacy and security of personal data. As an example, no personal data should be collected without the permission of the \emph{data subject} (i.e. user), stated in GDPR Article 13 and 14 \cite{noauthor_guide_nodate}.

It is important to understand what a DT is and what it is not. We should try to "close the loop" by sending informed decisions back to the physical counterpart so that it reflects the newest state of the DT. A DS could be used alongside the DT with the only goal of being a timestamped snapshot of some specific physical asset \cite{bergs_concept_2021}. When implementing one should also think about the challenges of digital twins, such as privacy concerns and costs. We should not get too caught up with modeling an exact 3D replica of an asset. On the other hand, we \emph{should} use existing semantic technologies such as knowledge graphs for formal knowledge representation from a variety of data sources.

\subsection{Semantic Web}
Semantic Web was \emph{envisioned} by Sir Tim Berners-Lee as an extension of the World Wide Web (WWW) in which data was linked. Linked data is about making links so people and machines can explore the web of data, and according to Berners-Lee, should follow the four rules \cite{tim_berners-lee_linked_nodate} below:

@@

    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.16]{graphics/rdf_triple_simple.png}
        \caption{Example of an RDF triple consisting of a subject, a predicate, and an object.}
        \label{fig:rdf_triple_simple}
    \end{figure}

@@

\begin{itemize}
    \item \textbf{Web Ontology Language (OWL)} is a language for expressing rich knowledge of concepts (phrases in natural language) in an application domain \cite{szolovits_overview_1977}, as well as relations between entities. More specifically, ontologies describe domains in terms of classes, properties, and individuals \cite{bechhofer_owl_2009}.

@@

There also exist other technologies that provide reasoner tools. Some notable mentions include, but are not limited to:
\begin{itemize}
    \item{\textbf{Apache Jena}} is an open source framework that has the Inference API which include reasoner tools that can be used to reason over data, as well as checking content of triple stores \cite{noauthor_apache_nodate-1}. The source code is publicly available\footnote{\url{https://github.com/apache/jena}}.
    \item{\textbf{OWL API}} is a Java framework that lets us create, serialize or manipulate ontologies written in OWL \cite{noauthor_owl_nodate}, and its source code is openly accessible\footnote{\url{https://github.com/owlcs/owlapi}}. Its recent development is more focused towards OWL 2 however.
    \item\hyperref[subsec:SMOL]{\textbf{SMOL}} provide reasoning tools through its interpreter, which implements \emph{semantic lifting} (i.e. generating a knowledge graph from the current state of a SMOL program) \cite{noauthor_semantic_nodate}.

@@

\subsection{Databases}\label{subsec:Databases}
Databases are collections of information that exist for a long time, often over many years \cite{garcia-molina_database_2002}. They can be divided into relational and non-relational (NoSQL) databases and are often compared as such \cite{mohamed_relational_2014}. Relational databases have existed for many decades and store data in tables (which consist of columns and rows), whilst \emph{NoSQL} (i.e. not only SQL) databases store data in JSON documents \cite{garcia-molina_database_2002, sudiartha_data_2020}. Examples of relational databases are MySQL and PostgreSQL, whereas examples of NoSQL databases are: 
\begin{itemize}
    \item \textbf{MongoDB} is a well-known NoSQL database for storing data in JSON documents and its source code is publicly available\footnote{\url{https://github.com/mongodb/mongo}}.
    \item \textbf{Firebase Realtime Database} is a cloud-hosted NoSQL database that stores data as JSON objects  \cite{noauthor_firebase_nodate}. Furthermore, data is synchronized in real-time to cross-platform clients.
    \item \textbf{InfluxDB} is a NoSQL database for storing \emph{time series data} (i.e. "sequence of data points indexed in time order" \cite{noauthor_what_nodate})  \cite{noauthor_influxdb_nodate}, and its source code is also openly accessible\footnote{\url{https://github.com/influxdata/influxdb}}.
\end{itemize}

@@

\newpage
\section{Problem analysis and design}\label{sec:Analysis}
This chapter examines the problem statement (H) and its three guiding research questions (RQ1, RQ2, RQ3) in \ref{sec:Introduction}(Introduction), based on the theoretical background information and general code examples in \ref{sec:Background}(Background). We create a building domain based on the analysis of the research questions. Then we examine what a dynamic asset model entails, and how static data can be separated from dynamic data in it. Different technologies described in \ref{sec:Background}, such as databases, frameworks for creating applications (apps), and semantic digital twins are considered. Lastly, we present the resulting structural requirements as the basis for the implementation.

@@

\subsubsection{Second research question (RQ2)}\label{subsubsec:RQ2}
RQ2 asks whether we can "enable bidirectional data flow between the PT and DT, such that the digital twin gets updated sensor data from the mobile devices in the PT and sends informed decisions back". Bidirectional data flow is data that is exchanged both ways, and not just in one direction between the physical twin (PT) and digital twin (DT), or vice versa. The DT observes the PT in near real-time and gets updated sensor data from its physical counterpart. The DT should then send informed decisions back to it. A lot of value is derived from the DT making an informed decision on behalf of a mobile asset in the PT, and it is crucial that the appropriate smartphone user is notified.

The DT could conclude to make an informed decision if some criteria are met. As an example of a criteria, we look at how we can check if a smartphone is inside a critical area. Using what is shown in Figure \ref{fig:simple_building} as a basis, an area's two border corners are expressed as follows:
\begin{enumerate}
    \item $(x_1, y_1)$
    \item $(x_2, y_2)$
\end{enumerate}
Meaning, if we want to check if a smartphone's physical location (point) $(x, y)$ is inside such an area, we can check if the following is true:\newline\newline$(x >= x_1\:AND\: x <= x_2)\:OR\:(x >= x_2\:AND\:x <= x_1)$\newline$AND$\newline$(y >= y_1\:AND\:y <= y_2)\:OR\:(y >= y_2\:AND\:y <= y_1)$\include\newline We also include both directions so maintainers can freely choose border corners for an area, which makes it easier to use and aligns with the RQ1.

@@

\subsection{Building domain}
From considering the three research questions, we have created a possible domain including a built environment based on static data, as well as movable entities with their physical location as dynamic data, as shown in Figure \ref{fig:static_built_environment}. \textbf{Physical 2D space} is provided for context of outside and is in two dimensions as described in \ref{subsec:Scope}(Thesis scope and outline). This means that altitude is not taken into account. Note that the relationships between some entities use \emph{composition} (i.e. indicates a very strong relationship between the entities, which means that if the owing entity (e.g. \textbf{Physical 2D space}) is destroyed, so is the entity linked to it (e.g. \textbf{Building} \cite{pilone_uml_2005}) (it can be thought of as a black hole..)

@@

It is important to look at the differences between the sensor data and the informed decisions in terms of what the data can be used for. The DT gets updated sensor data from the PT, and sends informed decisions back, as RQ2 states. The updated sensor data from the PT is the newest physical location from the mobile assets in the PT. It is crucial that the PT can use this data \emph{as is}, and therefore the data should be exact coordinates (latitude, longitude), and identifiable. We should think about what an informed decision means. An informed decision is a conclusion made by the DT when there is sufficient information to determine something on behalf of the PT to control it. An example of this is the DT sending warnings back to the PT when the DT has determined that smartphones are inside critical areas. The analysis of what a critical area \emph{is} however, is not taken into account, as stated in \ref{subsec:Scope}(Thesis scope and outline). Therefore, the DT should not decide that an area is dangerous. This could be a job for an operator responsible for defining the areas in the asset model.

@@

The \textbf{Server} in Figure \ref{fig:initial_components} has not been explicitly described yet in this thesis as its main function was to enable a bidirectional flow of data between the PT and the digital twin (DT). Most people in general are familiar with Java,  a well-known and established programming language, and our fellow students are quite familiar with its syntax as well. Lastly, The OWL API and Apache Jena have many open-source examples in Java, which is another reason we should use Java in our server implementation and not e.g. Kotlin (which the SMOL interpreter is written in).

\subsubsection{Semantic Digital Twins}\label{subsubsec:SemanticDigitalTwins}
\ref{subsec:SMOL}(SMOL) describes that SMOL (Semantic Micro Object Language) can be used as a framework for creating digital twins (DTs) and that it integrates semantic technologies. Knowledge graphs can be generated by using the \verb|dump|-command in the REPL provided by the SMOL interpreter. The knowledge graph can capture a knowledge base, such as a dynamic asset model, in which static data is separated from dynamic data. This asset model combined with the SMOL program, produces the knowledge graph. It is also possible to access the asset model from the SMOL program. As SMOL also integrates semantic languages, such as SHACL for validation and SPARQL for querying, it could be a good fit for our implementation.

@@

\noindent\textbf{Structural requirements:}
\begin{itemize}
    \item \textbf{Client}
    \begin{itemize}
        \item Client can choose to send its newest status (ID, latitude, longitude) 
        \item Appropriate endangered clients need to be warned 
        \item Show appropriate in-app messages to the user
        \item Ask for permission to track the physical position of the device
        \item Clearly show when the physical location of the device is tracked
        \item Clearly show when the physical location of the device is \emph{not} tracked
    \end{itemize}
    \item \textbf{Server}
    \begin{itemize}
        \item \textbf{1.} Forward the sensor data to a time series database, which the DT accesses
        \item \textbf{2.} Forward the informed decisions to a real-time database, which the PT accesses
        \item Should be possible to do \textbf{1.} and \textbf{2.} above at the same time
        \item Reloads asset model with tracked clients
        \item Access and read the generated knowledge graph
    \end{itemize}
    \item \textbf{Databases}
    \begin{itemize}
        \item Realtime database
        \begin{itemize}
            \item Store the ID and physical location of the clients
            \item Store data about which clients (in the PT) the DT made an informed decision on behalf of
        \end{itemize}
        \item Time series database
        \begin{itemize}
            \item Store the time series data
        \end{itemize}
    \end{itemize}
    \item \textbf{Semantic Digital Twin}
    \begin{itemize} 
        \item \textbf{3.} DS gets the newest sensor data from the time series database
        \item \textbf{4.} Access the smartphones by their IDs in the asset model
        \item Create smartphone objects by combining data from \textbf{3.} and \textbf{4.} above. 
        \item Access \emph{only} the areas from the asset model that are \emph{critical}
        \item Check if smartphones are inside any critical areas
        \item Handle if there are any smartphones that don't have a registered physical location
        \item Provide examples of SPARQL queries
        \item Show informative messages in the REPL during program execution
    \end{itemize}
\end{itemize}

@@

\section{Implementation}\label{sec:Implementation}
This chapter describes the proof-of-concept implementation based on the structural requirements in \ref{subsec:Requirements}(Structural requirements), by describing each of the following parts in it:
\begin{itemize}
    \item Client
    \item Server
    \item Databases
    \item Semantic Digital Twin
\end{itemize}
In the first section we describe an \emph{early version} of the final implementation based on the initial architecture of software components in Figure \ref{fig:initial_components}.

Then we look at the revised final version in Figure \ref{fig:components}, and describe what we implemented. For each of the larger architectural components in the figure, we reference the section that describes it in \ref{sec:Background}(Background), and the design decision that enables it in \ref{sec:Analysis}(Analysis and design). We also reference the corresponding \emph{item} in the structural requirement \emph{part} we are in.

\subsection{Early implementation}\label{subsec:EarlyImplementation}
Due to the extensive architecture, we chose to implement an early version by letting all of the architectural components in Figure \ref{fig:initial_components} communicate first. The \textbf{Physical Twin} in the figure consists of \textbf{Mobile Assets} which are smartphones. Based on the background information in \ref{subsec:AppDevelopment}(Background - App Development), and from further narrowing of the technologies in \ref{subsubsection:AppDevelopment}(Analysis - App Development), a Flutter app was created for cross-platform support.

Figure \ref{fig:recording_off} and \ref{fig:recording_on} show the location recording toggle function in the app, off and on, respectively. The recording is off by default, and when the button is pressed, it starts tracking the position of the device and changes color. Then when it is pressed again, the latest physical location of the device is sent, alongside its ID, as a status message to the server over TCP (Transmission Control Protocol).

@@

We created a Maven project (a tool for managing the project) and used the object-oriented programming language Java for the server implementation. To let the server track many clients, we first assigned an ID to the client when it connected to the server, and kept track of all assigned clients. The client received its assigned ID from the server which was later sent alongside the client's physical location as status update messages when pressing the button again. As this was an early implementation, we implemented client-server communication over TCP (Transmission Control Protocol) which is further discussed in \ref{subsec:ClientServerCommunication}(Client-server communication).

An asset model (ontology) of a simple building was also created as shown in Figure \ref{fig:simple_building} (this time with coordinates as data properties which were "added" to individuals of the \verb|MovableEntity| class suggested in \ref{subsubsec:DynamicAssetModel}(Dynamic asset model). We used Protégé to create the domain and exported it as an OWL document (file). This file was then accessed in a new SMOL program based on the background information in \ref{subsubsec:Technologies}(Technologies) and \ref{subsubsec:CoreLanguage}, and from further analysis in \ref{subsubsec:StaticAndDynamicData}(Static and dynamic data).

@@

\subsection{Implemented solution}\label{subsec:ImplementedSolution}
Figure \ref{fig:components} shows the architectural components in the final implementation. More specifically, it illustrates the \textbf{Mobile Assets} in the \textbf{Physical Twin}, its \textbf{Digital Twin} based on a dynamic \textbf{Asset Model}, and the \textbf{Server} which enables the communication between the PT and DT both ways. In addition to this, the figure illustrates that \textbf{Firebase Realtime database} is used as a channel between the \textbf{Physical Twin} and the \textbf{Server}, and that \textbf{InfluxDB time-series database} is used for storing the sensor data which \emph{can} be accessed from the \textbf{SMOL} program, as shown in the example function in \ref{fig:access_time_series}.

@@

In the following sections, we describe the architectural \emph{components} in the figure one by one based on the background information and the analysis, and by referencing the corresponding \emph{part} (Client, Server, Databases, or Semantic Digital Twin) in the structural requirements.

For further reference on how we setup SMOL to be used in this thesis, see \ref{subsec:Setup}(Setup).

\subsubsection{Client}
A Flutter app was written in Dart, based on the background information on app development in industry in \ref{subsec:AppDevelopment}(Background - App Development), and from the analysis of related technologies in \ref{subsubsection:AppDevelopment}(Analysis - App Development) which describes that Flutter is popular in industry. \ref{subsec:EarlyImplementation}(Early implementation) describes the toggling of location recording in the app, but from the paragraph about GDPR (General Data Protection Regulation) in \ref{subsec:DigitalTwins} and the requirement "Ask for permission to track the physical position of the device" we implemented location permissions. The section \ref{subsec:privacy}(Privacy) discusses privacy further.

The source code of the app was structured into the following parts: 
\begin{itemize}
    \item \verb|main.dart| which initializes the app and Firebase
    \item \verb|map.dart| which contains code for the UI (User interface) in the app. This is done by rendering Flutter widgets that behave much like components in e.g. React. Simply put, the UI is what the user \emph{sees} in the app and is built from the widgets.
    \item \verb|service.dart| accesses Firebase, and writes to, or reads from it.
    \item the model-class \verb|status.dart| is used to create status objects which are sent as JSON to Firebase in which it is stored as a JSON object, as described in \ref{subsec:Databases}(Background - Databases).
\end{itemize}
We used the Google Maps SDK (Software Development Kit) in the implementation to integrate the map and its functions in the app, as described in \ref{subsec:AppDevelopment}(App Development).

Some of the requirements set forth in the structural requirements were already implemented in the earlier version of the app, such as "Client can choose to send its newest status (ID, latitude, longitude)" However, there are two requirements stating that we have to \emph{clearly} show when the recording function is off/on. In order to clearly show it for those that are color blind as well, the colors of the button changed to blue and red, respectively. Figure \ref{fig:color_theme} shows the final color theme, which is located in the previously mentioned file \verb|main.dart| and can be accessed from anywhere in the app.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.50]{graphics/color_theme.png}
    \caption{The color theme in the app.}
    \label{fig:color_theme}
\end{figure}
